---
title: "R Bridge Final"
author: "Ann Liu-Ferrara"
date: "August 8, 2016"
output: html_document
---

This is R Bridage Final project. Data are a subset of hflights. Visualizing the airline arrival delay and depature delay by carriers, and exploring the relationships delays with other factors.

```{r}
library(hflights)
library(ggplot2)
```
Subsetting data and displaying descriptive summary statistics
```{r}
sub.delay <- hflights[hflights$Month == 12, 
                c("ArrDelay", "DepDelay", "UniqueCarrier", "DayOfWeek",
                  "AirTime", "Year", "Month", "DepTime", "Distance")]

str(sub.delay)
head(sub.delay)
summary(sub.delay)
```
Facet the delays by Day of Week
```{r}
ggplot(sub.delay, aes(x = as.factor(ArrDelay))) + geom_bar() + 
  coord_cartesian(xlim = c(0, 200)) +
  facet_wrap(~DayOfWeek, nrow = 1) +
  xlab("Arrival Delay") + 
  ylab("frequency")+
  labs(title = "Arrival Delay by Day of the Week in Dec. of 2011") 

ggplot(sub.delay, aes(x = as.factor(DepDelay))) + geom_bar() + 
  coord_cartesian(xlim = c(0, 200))  + 
  facet_wrap(~DayOfWeek, nrow = 1) +
  xlab("Departure Delay") + 
  ylab("frequency")+
  labs(title = "Depature Delay by Day of the Week in Dec. of 2011") 

```
Rank carriers by delay frequence and explore relationships between delays and distance.
```{r}
ggplot(sub.delay[, c("ArrDelay", "UniqueCarrier")], 
       aes(x=reorder(UniqueCarrier, UniqueCarrier,
                         function(x)-length(x)))) + geom_bar() +
  xlab("Arrival Delay by Carrier") + 
  ylab("Arrival Delay")+
  labs(title = "Arrival Delay by Carriers in Dec. of 2011") 
  
ggplot(sub.delay[, c("DepDelay", "UniqueCarrier")], 
       aes(x=reorder(UniqueCarrier, UniqueCarrier,
                     function(x)-length(x)))) + geom_bar() +
  xlab("Depature Delay") + 
  ylab("Departure Delay") +
  labs(title = "Depature Delay by Carriers in Dec. of 2011") 

ggplot(sub.delay[sub.delay$ArrDelay > 0, c("ArrDelay", "DepDelay")],
       aes(x = DepDelay, y = ArrDelay)) + geom_point() +
       xlab("Depature Delay") + 
       ylab("Arrival Delay") +
       labs(title = "Depature Delay vs. Arrival Delay in Dec. of 2011") 

ggplot(sub.delay[sub.delay$ArrDelay > 0, c("ArrDelay", "Distance")],
       aes(x = Distance, y = ArrDelay)) + geom_point()+
  xlab("Distance") + 
  ylab("Arrival Delay") +
  labs(title = "Distance vs. Arrival Delay in Dec. of 2011") 

```

The observations show:
  - Delay data have outliers.
  - Departure delay leads to arrival delay.
  - Delays are the highest on Friday
  - The carriers on the top frequency of delay list: XE, CO, WN, and OO.
  - Distance has no direct impact on delays.